Sampling Strategies in Decisions from Experience

### 2021-12-09 manuscript.Rmd
Review of Introduction

Now, consider the power set $\cal{P}(A)$, which is the set of all possible subsets of $A$.
Each of the subsets $A' \in \cal{P}(A)$ corresponds to a possible distinction between choice alternatives that can be made after assessing for each alternative whether it fulfills the selection criteria or not [@jostMathematicalConcepts2015, p. 16]. 
The concept of the power set thereby illustrates in a primitive way how different selection criteria can affect the consequential choice, provided the properties that can be captured and manipulated by a particular processing system, i.e., a living organism or a computer, are not uniformly distributed among the alternatives.

@@

A rather important part---though by far not the only one---of the computation that connects the initial set with its subset is the comparison of the properties of the choice alternatives against some selection criteria. 
At the algorithmic level [cf. @marrVisionComputationalInvestigation1982], such comparisons require an explicit description of the properties using a representation system which is useful to the processing system, i.e., a living organism or a computer [@marrRepresentationRecognitionSpatial1978]. 
In quite many cases, however, the selection process is not restricted to one particular representation system but a number of different representations can be used to describe the properties of choice alternatives in a given set.
Yet, not every representation system permits the description of a given property.
E.g., a consequential difference that has emerged in behavioral decision research is whether the probability of a risky outcome is readily described, i.e., as a decimal number (.2), ratio (2/10), or percentage (20%), or whether it must be inferred from the relative frequency of occurrence of the particular outcome in a collection of i.i.d. random variables [@hertwigDescriptionexperienceGapRisky2009]. 
Whereas all four representation systems can in principal be used to describe the probabilities of given outcomes, only the representation which uses collections of random variables can, for example, plausibly produce wavy recency effects due to the perception of spurious temporal outcome patterns [@plonskyRelianceSmallSamples2015; @plonskyLearningSettingsPartial2017].
Although a discrete-time collection of i.i.d. random variables cannot, by definition, exhibit temporal regularity in the sequence of sampled outcomes, spurious temporal patterns might nevertheless be regarded as a property of choice alternatives and affect the consequential decision behavior in a way that would not be possible in other representation systems. 
The reason is simply that such a (spurious) property could not emerge in the other representations.

The preceding example illustrates how some properties can be described in one representation but not in another and how representation systems can differ with regards to the processing effort required to describe the same property.
Thus, there is generally a trade-off between different representation systems, where in a given representation certain properties become explicit while others are pushed into the background [@marrVisionComputationalInvestigation1982, p. 21].

>>>>

Now, consider the power set $\cal{P}(A)$, which is the set of all possible subsets of $A$.
Each of the subsets $A' \in \cal{P}(A)$ corresponds to a possible distinction between choice alternatives that can be made after assessing for each alternative whether it fulfills the selection criteria or not [cf. @jostMathematicalConcepts2015, p. 16]. 
The concept of the power set thereby illustrates in a primitive way how different selection criteria can affect the consequential choice, provided the properties that can be captured and manipulated by a particular processing system, i.e., a living organism or a computer, are not uniformly distributed among the alternatives.

@@

A rather important part---though by far not the only one---of the computation that connects the initial set with its subset is then the comparison of the properties of the choice alternatives against some selection criteria. 
At the algorithmic level, such a comparison requires an explicit description of the properties using a *representation* which is useful to the processing system [cf. @marrVisionComputationalInvestigation1982]. 
In quite many cases, however, a number of different representations can be used to describe the properties of choice alternatives in a given set.
Yet, not every representation permits an accurate description of a given property at reasonable expenses or allows to describe a property at all [cf. @marrRepresentationRecognitionSpatial1978].
E.g., a consequential difference that has emerged in behavioral research on risky choice is whether the probability of a risky outcome is described using a decimal number (.2), ratio (2/10), or percentage (20%), or whether it must be inferred from the relative frequency of occurrence of the particular outcome in a collection of independent and identically distributed (i.i.d.) random variables, i.e., a *random sample*.
Noteworthy, for the latter case, which is usually referred to as *decisions from experience*, empirical protocols indicate that people decide as if they were underweighting small probability events [@barronSmallFeedbackbasedDecisions2003; @hertwigDecisionsExperienceEffect2004; @weberPredictingRiskSensitivity2004].
One undisputed factor that contributes to this underweighting pattern is the sampling error associated with people's reliance on small samples [e.g., @foxDecisionsExperienceSampling2006; @hertwigDecisionsExperienceEffect2004; @hauDescriptionexperienceGapRisky2008; @ungemachAreProbabilitiesOverweighted2009; see also @wulffMetaanalyticReviewTwo2018 for a meta-analytic review], indicating that, although the probabilities of outcomes can in principle be accurately described in decisions from experience, the actual description is rather inaccurate because it cannot be done at reasonable expenses.

Another property of choice alternatives that is likely to play a role in risky choice but can only be made explicit using a sample representation as in decisions from experience is the sequence in which outcomes occur.
Although a discrete-time collection of i.i.d. random variables cannot, by definition, exhibit temporal regularity, i.e., it cannot produce a stable sequence of sampled outcomes, spurious temporal patterns might nevertheless be regarded as a property of choice alternatives and affect the consequential choice behavior. 
Indeed, the underweighting pattern and other behavioral phenomena like the wavy recency effect and probability matching, are reasonably well captured by models which incorporate the assumption that people rely on (spurious) outcome sequences [@cohenEffectPerceivedPatterns2021; @gaissmaierSmartPotentialProbability2008; @plonskyRelianceSmallSamples2015; @plonskyLearningSettingsPartial2017;  @yuSequentialEffectsSupersition2009].

The preceding two examples illustrate how some properties of choice alternatives can be described in one representation but not in another and how representations can differ with regards to the processing effort required to describe the same property.
Thus, there is generally a trade-off between different representations, where in a given representation certain properties become explicit while others are pushed into the background [@marrVisionComputationalInvestigation1982, p. 21].
